Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-07-14 doc/report/msc-report.tex
report: method.

@@

@@

% TODO
The first section in this chapter will present the work procedure during the
master thesis. The following sections will discuss selected tools and their
impact on design and implementation choices for the Zohmg system. The last
section will describe the scope of the thesis.

@@

% TODO: generally use passive tense.
% TODO: renskriv.

This section describes how we structured our work over the course of the thesis
project. It describes how we, by working in an exploratory fashion, reached a
good understanding of the problem at hand, and how we successively moved closer
to a working solution by building early prototypes and iteratively improving
upon them.

The design phase, which we entered once our understanding of the problem scope
and its theoretical solution was satisfactory, and the following implementation
are described, followed by a summary of our tools studies.

\subsection*{Beginning}

% not so sure I like this section much..

We entered this project with a basic understanding of the problem Last.fm wanted
to solve but with little formal knowledge of the problem domain. We were tasked
with building a data store that would hopefully replace the store that was
presently in use.

We met the relevant people, were introduced to the current system and were
presented with literature in the area, namely OLAP Solutions and Whatever in
Data Warehousing.

\subsection*{Exploration}

The possible tools were explored by testing and benchmarking them.

The problem area was explored by prototyping solutions to a given subproblem.

\subsubsection*{Software}

@@

\subsubsection*{Benchmarks}

Only HBase made it this far, the other databases in the roster were hard or
impossible to either compile, install, setup, or all of them.

% TODO: need data on this.
HBase was benchmarked (the benchmarks was similar to the ones found in the
BigTable paper?).

\subsubsection*{Prototyping}

The problem domain was explored by asking, the system to be built and the
requirements upon it.

To explore the environment, for example, we asked what data sources there are
(web logs, scrobbling logs and radio streaming logs) and where that data is
stored (on HDFS, it turned out). We asked how much data the system was supposed
to handle (lots) and what the output format ought to be (any format, within
reason).

Most of the exploration was straight forward: many questions could be answered
from observations or by logical reasoning. Others were more open-ended, such as
deciding on a suitable output format. For the open-ended questions we would list
reasonable options and pitt them against each other, first on the drawing board
and eventually as prototypes.

% this is a little thin. The prototypes are small pieces of code that would
TODO: answer questions such as Is it at all possible to.. ? and What is the most
efficient way to..?

\subsection*{Literature studies}
% vet inte riktigt vad jag vill saga om det har.

Studying previous work in related fields brought OLAP systems to our attention.
It was to become very influential in our theoretical understanding of the
subject.

\subsection*{Python}

@@

After having become familiar with the problem and its theoretical solution, it
was time to design the software.

@@

The implementation phase was rather small: we would simply stitch together our evolved prototypes and arrive at a solid ground to stand on: we had a system for running MapReduce jobs, storing data in HBase, etc.

Writing code on top of this that performed data cube manipulation was a kinch!

\section{Apache Hadoop}

@@

\subsection*{Hadoop Distributed File System}

@@

\subsection*{Hadoop streaming}

@@

\subsubsection*{Dumbo}

@@

\section{Apache HBase}

@@

\subsection*{Physical storage}

@@

\subsection*{Sparsity}

@@

\subsection*{Hierarchies}

@@

\subsection*{Thrift}

@@

\section{Dependability}

@@

Because BigTable stores data sorted on row keys, it was decided that Zohmg only
supports time-series-based data. Having data sorted between nodes in a computer
cluster balances the network, computational, and storage loads, resulting in
more even utilization of the cluster. Since Zohmg only supports time-series,
each query has to be made over a time range.

@@

% TODO
The first section of this chapter talked about the procedure during the thesis
work. It described the exploratory model and the prototyping performed, as well
as the design phase that led up to the implementation.

The following sections presented and discussed the selected tools; they also
discussed the impact on design and implementation choices.

The last section describes the scope of the thesis.

>>>>

TODO:
Other techniques that have been frequently used to build data warehouses
include: A, B, and C, amongst others.

A solves distributed computing by ...
B is blablabla ...
The framework C was ...

@@

TODO: expand.

@@

The first section describes how the thesis project work was begun and later
structured. The design and implementation phases are also described in this
section.

The following two sections presents the tools which were selected to build Zohmg
upon.

The last section presents the scope of the thesis.

@@

This project started with an introduction to existing tools and infrastructure
at Last.fm. The introduction also included specifying requirements on the final
product, and introduction to several important concepts in data warehousing
(such as OLAP). The concepts were presented with related literature and research
papers.

The general procedure of the thesis project was done as exploratory work, which
involved rapid prototyping on sub problems and iteratively improving these
prototypes. A working system solution successively moved closer when more of
these sub problems was solved in a satisfactory manner.

Once a thourough understanding of the problem scope and its theoretical solutian
the design were entered. Finally a short implementation phase was executed.

\subsection*{Literature Studies}

To establish good understanding of the problem in general and new concepts,
literature and research papers were studied. The concepts studied are data
warehousing, OLAP, distributed storage with column-oriented databases such
as BigTable, and distributed computing with the MapReduce pattern and
frameworks.

Also a lot of related and surronding, possibly useful, concepts were studied;
although some of these never made it to the final solution.
\cite{pig,mapreducemerge}

\subsection*{Software}

@@

\subsubsection{Benchmarks}

In order to evaluate software which candidated for taking responsibility for the
data storage, benchmarked were planned. Finally, only HBase made it to the
benchmarking part, the other databases in the roster were difficult or
impossible to either compile, install, setup, or all of them.

TODO: HBase was benchmarked (the benchmarks was similar to the ones found in the
BigTable paper?). NEEDS DATA!!

\subsection*{Prototyping}

As mentioned above, the problem domain was explored by prototyping solutions to
sub problems of the greater system. The prototypes were for example used to find
out if a problem was solvable at all, or what the most efficient implementation
of an idea was.

Most of the exploration was straight forward. Many questions could be answered
from observations or by logical reasoning, other questions were more open ended.
Such an open ended question was deciding on a suitable output format\footnote{An
output format is a concept in Hadoop; it is a format which the reduce step uses
to output the results.}. For open ended questions the approach was to first
reason about suitable solutions, compare them against each other on the drawing
board and then eventually as prototypes.

The first prototypes handled data sources. The different data sources had
different formats and locations, the related prototypes dealt with how these
data sources should be analyzed and imported with MapReduce. The succeeding
prototypes explored export and data storage functionality.

\subsubsection{Python}

@@

TODO: ALL OF THIS.
After having become familiar with the problem and its theoretical solution, it
was time to design the software.

@@

The implementation phase was rather small. The already existing evolved
prototypes, which solved sub problems, was simply stitched together. The result
gave a solid ground to stand on: A system which could run import programs
(MapReduce jobs), shuffle data from the import programs into HBase, and export
data from HBase through a HTTP API. Data cube implementation was added to this
resulting system.

\section{Tools}

\subsection{Apache Hadoop}

@@

\subsubsection{Hadoop Distributed File System}

@@

\subsubsection{Hadoop streaming}

@@

\subsubsection{Dumbo}

@@

\subsection{Apache HBase}

@@

\subsubsection{Physical storage}

@@

\subsubsection{Sparsity}

@@

\subsubsection{Hierarchies}

@@

\subsubsection{Thrift}

@@

\subsection{Dependability}

@@

Because BigTable stores data sorted on row keys, it was decided that Zohmg only
supports time-series-based data. Having data sorted between nodes in a computer
cluster balances the network, computational, and storage loads, resulting in
more even utilization of the cluster. Time-series analysis is also the most
common for logs, since each entry is time-stamped. \cite{discoveringweb} Since
Zohmg only supports time-series, each query has to be made over a time range.

@@

The first section of this chapter talked about the procedure during the thesis
work. It described the exploratory model and the prototyping performed, as well
as the design phase which led to the implementation.

The following sections presented and discussed the selected tools; they also
discussed the tools' impact on design and implementation choices.

The last section described the scope of the thesis.
